---
title: "Figures for the ggdist paper"
author: "Matthew Kay"
date: "2023-03-11"
output: html_document
---

## Introduction

This document generates the various figures for the paper, *ggdist: Visualizations of Distributions and Uncertainty in the Grammar of Graphics*.

## Setup

These libraries are needed to build this file:

```{r setup}
library(ggplot2)
library(ggdist)              # obviously
library(distributional)      # for dist_...
library(dplyr)               # for data manipulation
library(systemfonts)         # for using system fonts
```

And this ggplot theme, based on `theme_ggdist()`, using the Source Sans Pro font (in the `fonts/` folder):

```{r}
theme_set(
  theme_light(base_family = "Source Sans Pro") + theme(
    axis.line.x = element_line(color = "gray70", linewidth = rel(0.5)), 
    axis.line.y = element_line(color = "gray70", linewidth = rel(0.5)),
    axis.title.x = element_text(margin = margin(t = 7)), 
    axis.title.y = element_text(margin = margin(r = 7)), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.border = element_blank(),
    strip.text = element_text(color = "black", margin = margin(6, 6, 6, 6)),
    strip.background = element_rect(fill = "gray90")
  )
)
```

And some functions we'll use for annotating stuff:

```{r}
red_ = "#d95f02"
label_color = red_

bracket_ = function(..., x, xend = x, y, yend = y, color = label_color) {
  annotate("segment",
    arrow = arrow(angle = 90, ends = "both", length = unit(3, "points")),
    color = color, linewidth = 0.75,
    x = x, xend = xend, y = y, yend = yend,
    ...
  )
}
arrow_ = function(..., curvature = 0, x, xend = x, y, yend = y, color = label_color) {
  annotate("curve", 
    color = color, arrow = arrow(angle = 45, length = unit(3, "points"), type = "closed"),
    curvature = curvature,
    x = x, xend = xend, y = y, yend = yend
  )
}
label_ = function(..., hjust = 0, color = label_color) {
  annotate("text",
    color = color, hjust = hjust, lineheight = 1,
    size = 3.75,
    ...
  )
}
```



## Setting the stage

### A simplified notation for the grammar of graphics

We're going to use a slightly modified version of the mtcars dataset:

```{r cars}
cars = mtcars |>
  mutate(weight = wt)
```

We can generate a nice scatterplot of it:

```{r 2-mpg_v_weight, fig.height = 2, fig.width = 2.5}
p = cars |>
  ggplot(aes(x = weight, y = mpg)) +
  geom_point()
p

ggsave("figs/2-mpg_v_weight.pdf", device = cairo_pdf)
```

### Uncertainty as she is spoke

```{r 2-mean_sd_interval, fig.height = 1.25, fig.width = 2.5}
p = data.frame(a = dist_normal(0, 1)) |>
  ggplot(aes(xdist = a)) +
  stat_pointinterval(.width = .95) +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL, limits = c(-2.95, 2.95)) +
  label_(label = "italic(a)", x = 0, y = 0.19, hjust = 0.5, parse = TRUE) +
  bracket_(x = 0, xend = 1, y = -0.18) +
  label_(label = "italic(sigma[a])", x = 0.5, y = -0.34, hjust = 0.5, parse = TRUE) +
  labs(y = NULL, x = NULL)

p
ggsave("figs/2-mean_sd_interval.pdf", device = cairo_pdf)
```

```{r 2-xmin_xmax_interval, fig.height = 1.25, fig.width = 2.5}
p = data.frame(a = dist_normal(0, 1)) |>
  ggplot(aes(xdist = a)) +
  stat_pointinterval(.width = .95, justification = 0.55) +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL, limits = c(-2.95, 2.95)) +
  label_(label = "italic(a)", x = 0, y = 0.17, hjust = 0.5, parse = TRUE) +
  arrow_(x = -1.96, y = -0.2, yend = -0.06) +
  label_(label = "italic(a - 1.96%.%sigma[a])", x = -1.96, y = -0.32, hjust = 0.5, parse = TRUE) +
  arrow_(x = 1.96, y = -0.2, yend = -0.06) +
  label_(label = "italic(a + 1.96%.%sigma[a])", x = 1.96, y = -0.32, hjust = 0.5, parse = TRUE) +
  labs(y = NULL, x= NULL) 

p
ggsave("figs/2-xmin_xmax_interval.pdf", device = cairo_pdf)
```

## Distributional vis

### Intervals

```{r 3-geom_pointinterval_quantiles, fig.height = 1.5, fig.width = 2.5}
p = data.frame(a = dist_normal(0, 1)) |>
  ggplot(aes(xdist = a)) +
  stat_pointinterval(.width = .95, justification = 0.65) +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL, limits = c(-2.95, 2.95)) +
  label_(label = "median(italic(A))", x = 0, y = 0.15, hjust = 0.5, parse = TRUE) +
  arrow_(x = -1.96, y = -0.2, yend = -0.04) +
  label_(label = "{italic(F)[italic(A)]^{-1}}~bgroup('(', over(1 - gamma, 2), ')')", x = -1.96, y = -0.42, hjust = 0.5, parse = TRUE) +
  arrow_(x = 1.96, y = -0.2, yend = -0.04) +
  label_(label = "{italic(F)[italic(A)]^{-1}}~bgroup('(', over(1 + gamma, 2), ')')", x = 1.96, y = -0.42, hjust = 0.5, parse = TRUE) +
  labs(y = NULL, x = NULL)

p
ggsave("figs/3-geom_pointinterval_quantiles.pdf", device = cairo_pdf)
```

```{r 3-stat_pointinterval_linewidth, fig.height = 2.5, fig.width = 2.5}
p = data.frame(a = dist_normal(0, 1)) |>
  ggplot(aes(xdist = a)) +
  stat_pointinterval() +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL, limits = c(-2.5, 2.5)) +
  bracket_(x = -1.96, xend = 1.96, y = 0.07) +
  label_(label = "γ = 0.95", x = -1.72, y = 0.12, hjust = 0, vjust = 0) +
  bracket_(x = -0.95, xend = 0.95, y = -0.07) +
  label_(label = "γ = 0.66", x = -0.67, y = -0.12, hjust = 0, vjust = 1) +
  labs(y = NULL, x = NULL)

p
ggsave("figs/3-stat_pointinterval_linewidth.pdf", device = cairo_pdf)
```

```{r 3-stat_pointinterval_color, fig.height = 1.5, fig.width = 2.5}
p = data.frame(a = dist_normal(0, 1)) |>
  ggplot(aes(xdist = a)) +
  stat_pointinterval(aes(interval_color = after_stat(level)), linewidth = 18, .width = c(.5, .8, .95)) +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL, limits = c(-2.5, 2.5)) +
  scale_color_brewer(aesthetics = "interval_color") +
  labs(y = NULL, x = NULL, interval_color = "mass")

p
ggsave("figs/3-stat_pointinterval_color.pdf", device = cairo_pdf)
```


## Session Info

This file was built with the following setup:

```{r session info}
sessionInfo()
```
