---
title: "Figures for the ggdist paper"
author: "Matthew Kay"
date: "2023-03-11"
output: html_document
---

## Introduction

This document generates the various figures for the paper, *ggdist: Visualizations of Distributions and Uncertainty in the Grammar of Graphics*.

## Setup

These libraries are needed to build this file:

```{r setup}
library(ggplot2)
library(ggdist)              # obviously
library(distributional)      # for dist_...
library(dplyr)               # for data manipulation
library(systemfonts)         # for using system fonts
```

And this ggplot theme, based on `theme_ggdist()`, using the Source Sans Pro font (in the `fonts/` folder):

```{r}
theme_set(
  theme_light(base_family = "Source Sans Pro") + theme(
    axis.line.x = element_line(color = "gray70", linewidth = rel(0.5)), 
    axis.line.y = element_line(color = "gray70", linewidth = rel(0.5)),
    axis.title.x = element_text(margin = margin(t = 7)), 
    axis.title.y = element_text(margin = margin(r = 7)), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.border = element_blank(),
    strip.text = element_text(color = "black", margin = margin(6, 6, 6, 6)),
    strip.background = element_rect(fill = "gray90")
  )
)
```


## Setting the stage

### A simplified notation for the grammar of graphics

We're going to use a slightly modified version of the mtcars dataset:

```{r cars}
cars = mtcars |>
  mutate(weight = wt)
```

We can generate a nice scatterplot of it:

```{r fig.height = 2, fig.width = 2.5}
p = cars |>
  ggplot(aes(x = weight, y = mpg)) +
  geom_point()
p

ggsave("figs/2-mpg_v_weight.pdf", device = cairo_pdf)
```



## Session Info

This file was built with the following setup:

```{r session info}
sessionInfo()
```
