---
title: "Analysis of ggdist issues"
author: "Matthew Kay"
date: "2023-03-25"
output: html_document
---

## Introduction

This document does some simple data cleaning of issues exported from the `ggdist` Github repository for the paper, *ggdist: Visualizations of Distributions and Uncertainty in the Grammar of Graphics*.

## Setup

These libraries are needed to build this file:

```{r setup}
library(ggplot2)
library(ggdist)
library(jsonlite)
library(dplyr)               # for data manipulation
library(rlang)
library(readr)

theme_set(theme_ggdist())
```

## Data

The JSON file containing issues from the `ggdist` repository is `issues.json`, and
was generated using this [Github CLI](https://cli.github.com/) command on March 25, 2023:

```
gh issue list --json assignees,author,body,closed,closedAt,comments,createdAt,labels,milestone,number,id,title,reactionGroups,state,updatedAt,url --limit 1000 --state all > issues.json
```

We will load it into R:

```{r load_json}
issues = read_json("issues.json")
```

Let's turn this into a data frame at the top level so it's easier to work with:

```{r}
issues_df = bind_rows(lapply(issues, \(issue) {
  # some fields need to be wrapped in a list so the output is a list column
  if (length(issue$comments) > 0) issue$comments = list(issue$comments)
  issue$labels = list(vapply(issue$labels, `[[`, "name", FUN.VALUE = character(1)))
  issue$assignees = list(vapply(issue$assignees, `[[`, "login", FUN.VALUE = character(1)))
  # for some fields, we only want the human-readable name, not metadata
  issue$author = issue$author$login
  issue$milestone = issue$milestone$title
  # encode zero-length entries as NA
  issue = lapply(issue, \(entry) if (length(entry) == 0) NA else entry)

  issue_df = as_tibble(issue)
  if (nrow(issue_df) != 1) {
    print(issue$number)
    stop()
  }
  # stopifnot(nrow(issue_df) == 1)
  issue_df
})) |>
  mutate(
    closedAt = parse_datetime(closedAt),
    createdAt = parse_datetime(createdAt),
    updatedAt = parse_datetime(updatedAt),
  )

glimpse(issues_df)
```

Now, let's encode the remaining list-columns into strings so we can output something useful to CSV format:

```{r}
issues_df |>
  mutate(
    assignees = vapply(assignees, paste0, collapse = ",", FUN.VALUE = character(1)),
    labels = vapply(labels, paste0, collapse = ",", FUN.VALUE = character(1)),
    comments = sapply(comments, \(comments) {
      paste0(lapply(comments, \(comment) 
        paste0(
          "----------------------------\n", 
          comment$author$login, ":\n\n", 
          comment$body
        )),
      collapse = "\n\n")
    })
  ) |>
  write_excel_csv("issues.csv")
```

After this export, I manually imported the above CSV into Excel and tagged some issues (the `from_users` and `notes` columns) and saved the output in `issues_tagged.xlsx` and `issues_tagged.csv`. Let's import that updated csv now:

```{r}
issues_tagged = read_csv("issues_tagged.csv",
  col_type = cols(
    closed = col_logical(),
    number = col_double(),
    from_users = col_logical(),
    createdAt = col_datetime(),
    closedAt = col_datetime(),
    updatedAt = col_datetime(),
    .default = col_character()
  )
)
```

## Summaries

Closed issues:

```{r}
table(issues_tagged$closed)
prop.table(table(issues_tagged$closed))
```
Open issues are in these categories:

```{r}
table(unique(issues_tagged[!issues_tagged$closed,]$notes))
```

Number of authors (including me):

```{r}
length(unique(issues_tagged$author))
```

Number of issues created by users directly:

```{r}
table(issues_tagged$author != "mjskay")
prop.table(table(issues_tagged$author != "mjskay"))
```


Number of issues that stemmed from users, even if not created by them directly (e.g., because I created an issue based on a user's comment on another issue, or from a Twitter conversation with that user, etc):

```{r}
table(issues_tagged$from_users)
prop.table(table(issues_tagged$from_users))
```

Breakdown of categories: first, all unique combinations:

```{r}
table(issues_tagged$notes)
```

We don't really care about issues that are *purely* internal: e.g., refactoring, code cleanup, performance, etc. So we'll just lump those all together as "internal":

```{r}
categories = issues_tagged$notes
categories[categories %in% c("refactor", "cleanup", "internal", "performance")] = "internal"
```

Then we'll count categories (note: these are not unique per-issue):

```{r}
cat_counts = sort(table(unlist(strsplit(categories, ","))))
cat_counts / nrow(issues_tagged) #can't use prop table because categories overlap
```

